Modern BTree\\ techniques

### 2020-01-11 btree/presentation.tex
Normalized keys in indirection vector

\tikzset{rum-node/.style={draw,circle,fill=white,minimum width=2cm}}
\tikzset{rum-extra-node/.style={rectangle, draw, dashed, rounded corners, fill=white }}
\tikzset{btree-key/.style={minimum height=1cm, pattern=north west lines, draw}}
\tikzset{btree-pointer/.style={minimum height=1cm, draw}}
\tikzset{btree-hide/.style={draw opacity=0, line width=0, pattern=none}}
\tikzset{btree-empty/.style={btree-pointer, dashed}}
\tikzset{btree-leaf/.style={fill=greenGood}}
\tikzset{btree-branch/.style={fill=gray!20}}
\tikzset{btree-line/.style={line width=0.5pt}}
\tikzset{btree-path/.style={btree-pointer, minimum width=0.8cm, fill=red, opacity=0.5}}
\tikzset{>=latex}
\tikzset{btree-high-key/.style={btree-key, pattern=north west lines, pattern color=red, draw=red}}
\tikzset{btree-key-prefix/.style={btree-key, minimum height=0.2cm, fill=red, opacity=0.5}}
\tikzset{btree-key-compare/.style={btree-key, minimum height=0.6cm, fill=blue, opacity=0.5}}
\tikzset{arrow-pointer/.style={
        single arrow,
        minimum height=1.5cm,
        inner sep=3pt,
        line width=1pt,
        draw,
        color=gray,
        single arrow tip angle=45,
        single arrow head extend=0.1cm
    }
}

@@

\newcounter{itemid}
\newcounter{itemid-prev}

@@

@@

    \begin{overprint}[8cm]
    \end{overprint}

>>>>

\tikzset{rum-node/.style={draw,circle,fill=white,minimum width=2cm}}
\tikzset{rum-extra-node/.style={rectangle, draw, dashed, rounded corners, fill=white }}
\tikzset{btree-key/.style={minimum height=1cm, pattern=north west lines, draw}}
\tikzset{btree-pointer/.style={minimum height=1cm, draw}}
\tikzset{btree-hide/.style={draw opacity=0, line width=0, pattern=none}}
\tikzset{btree-empty/.style={btree-pointer, dashed}}
\tikzset{btree-leaf/.style={fill=greenGood}}
\tikzset{btree-branch/.style={fill=gray!20}}
\tikzset{btree-line/.style={line width=0.5pt}}
\tikzset{btree-path/.style={btree-pointer, minimum width=0.8cm, fill=red, opacity=0.5}}
\tikzset{>=latex}
\tikzset{btree-high-key/.style={btree-key, pattern=north west lines, pattern color=red, draw=red}}
\tikzset{btree-key-prefix/.style={btree-key, minimum height=0.2cm, fill=red, opacity=0.5}}
\tikzset{btree-key-compare/.style={btree-key, minimum height=0.6cm, fill=blue, opacity=0.5}}
\tikzset{arrow-pointer/.style={
        single arrow,
        minimum height=1.5cm,
        inner sep=3pt,
        line width=1pt,
        draw,
        color=gray,
        single arrow tip angle=45,
        single arrow head extend=0.1cm
    }
}
\tikzset{btree-indirection/.style={btree-key, minimum width=0.2cm, inner sep=0, pattern=none}}
\tikzset{btree-page-header/.style={btree-key, minimum width=0.5cm, inner sep=0, pattern=none}}

@@

\newcounter{itemid}
\newcounter{itemid-prev}
\newcounter{itemid-next}

@@

\newcommand{\btreenodewithindirectioninternal}[5] {
    \node[btree-page-header, #4] (page-header#1) {};
    \node[btree-indirection, #4, right=0.2cm of page-header#1] (indirection-key#11) {};
    \forloop{itemid}{2}{\value{itemid} < #2} {
      \setcounter{itemid-prev}{\value{itemid} - 1}
      \node[btree-indirection, right=0 of indirection-key#1\number\value{itemid-prev}]
        (indirection-key#1\number\value{itemid}) {};
    }

    \node[btree-key, #4, right=1cm of indirection-key#1\number\value{itemid-prev}] (key#11) {};
    \forloop[2]{itemid}{2}{\value{itemid} < #2} {
      \setcounter{itemid-prev}{\value{itemid} - 1}
      \setcounter{itemid-next}{\value{itemid} + 1}
      \node[btree-key, right=0.2cm of key#1\number\value{itemid-prev}, minimum width=1cm]
        (key#1\number\value{itemid}) {};
      \node[btree-key, right=0.2cm of key#1\number\value{itemid}] (key#1\number\value{itemid-next}) {};
    }
}

\newcommand{\btreenodewithindirection}[5] {
    \btreenodewithindirectioninternal{#1}{#2}{#3}{#4}{#5}
    \node[draw, inner sep=0.2cm, behind path, #5,
        fit=(page-header#1)
        \directlua{
          for index=1,#2 do
              tex.print("(key#1"..index..")")
          end}
        \directlua{
          for index=1,#2-1 do
              tex.print("(indirection-key#1"..index..")")
          end}
    ] (node#1) {};

    %% redraw after background
    \btreenodewithindirectioninternal{#1}{#2}{#3}{#4}{#5}
}

@@

    \begin{overprint}[8cm]
        \onslide<1>
        \begin{center}
        \begin{tikzpicture}[]
            \btreenodewithindirection{1}{5}{5}{}{btree-leaf}

            \foreach \i/ \shift in {1/0.35cm,2/0.11cm,3/-0.07cm,4/-0.3cm} {
                \draw[->, btree-line] ([yshift=\shift]indirection-key1\i.center) -- ([yshift=\shift]key1\i.center);
            }
        \end{tikzpicture}
        \end{center}

        \onslide<2>
        \begin{center}
        \begin{tikzpicture}[]
            \btreenodewithindirection{1}{5}{5}{}{btree-leaf}

            \foreach \i/ \shift/ \letter in {1/-0.6cm/a,2/-0.4cm/b,3/-0.2cm/c,4/-0.0cm/d} {
              \node[draw=none, below=\shift of indirection-key1\i.center]
                (normalized-indirection-key1\i) {\tiny{\letter}};
            }

            \foreach \i/ \shift in {1/0.35cm,2/0.11cm,3/-0.07cm,4/-0.3cm} {
                \draw[->, btree-line] ([yshift=\shift]indirection-key1\i.center) -- ([yshift=\shift]key1\i.center);
            }
        \end{tikzpicture}
        \end{center}

        \onslide<3>
        \begin{center}
        \begin{tikzpicture}[]
            \btreenodewithindirection{1}{5}{5}{}{btree-leaf}

            \foreach \i/ \shift/ \letter in {1/-0.6cm/a,2/-0.4cm/b,3/-0.2cm/c,4/-0.0cm/d} {
              \node[draw=none, below=\shift of indirection-key1\i.center]
                (normalized-indirection-key1\i) {\tiny{\letter}};
            }

            \foreach \i/ \shift in {1/0.35cm,2/0.11cm,3/-0.07cm,4/-0.3cm} {
                \draw[->, btree-line] ([yshift=\shift]indirection-key1\i.center) -- ([yshift=\shift]key1\i.center);
            }

          \linespread{0.5}
          \node[draw, below=1cm of indirection-key12.east, minimum width=1cm, minimum height=1cm, text width=1cm, align=center, rounded corners]
              (cpu-cache) {\footnotesize{CPU\\Cache}};

          \coordinate[above=2cm of cpu-cache.north east] (cpu-cache-top-left);
          \coordinate[above=2cm of cpu-cache.north west] (cpu-cache-top-right);
          \draw[dashed] ([xshift=-0.2cm]cpu-cache.north east) -- ([xshift=-0.2cm]cpu-cache-top-left);
          \draw[dashed] ([xshift=0.2cm]cpu-cache.north west) -- ([xshift=0.2cm]cpu-cache-top-right);

          \linespread{1.5}

        \end{tikzpicture}
        \end{center}
    \end{overprint}
